\vspace{-15mm}\fontsize{24pt}{10pt}\selectfont\textbf{A Radial Time Projection Chamber for $\alpha$ detection in CLAS at JLab}

### 2017-12-08 eg6-letters/NIM/EG6_RTPC_NIM.tex
update the nim paper with Stepan's comments.

\documentclass[preprint,5p]{elsarticle}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}  
\usepackage{amssymb}  
\usepackage{hyperref}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords:}} #1}
\hypersetup{colorlinks=true, urlcolor=blue, citecolor=blue}
\usepackage[displaymath]{lineno}
\linenumbers
\begin{document}

@@

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target extends along the beamline forming the detector 
     central axis; it is a 292~mm~long, 6~mm diameter Kapton straw with a 
     27~$\mu$m wall. The detector and the target are centered at 64~cm to the 
     left of CLAS's center.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 $\mu$m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~$\mu$m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The average electric field in this region is perpendicular to the beam 
      and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600~V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

The trigger signal also initiates RCU readout operation from FEC boards. All the 
measured samples from active channels are reported, together with a channel 
identifier and a timestamp, to the ROC application, that in turns sends them on 
the main event builder. Readout from FECs occurs in ``block-transfer'' mode, 
making use of the ALTRO internal multi-event buffer. This feature could be 
exploited thanks to the new RCU boards developed for the RTPC readout 
system\footnote{In the original BoNuS system, the U2F readout controller had no 
multi-event capability and used a USB connection rather than fiber optics for 
readout. These factors limited the maximum readout rate from the BoNuS detector 
to about 500 Hz.}. In order to read all the detector readout pads, four FEC 
crates were used, each equipped with 6 boards, plus a ROC. A schematic of the 
readout system, for a single crate, is reported in 
Figure~\ref{fig:ReadoutScheme}. This configuration permitted to reduce the dead 
time associated to FEC readout operations, which is scaling linearly with the 
number of boards in the crate. During the 2009 run, the system was successfully 
operated with a DAQ rate of 3.1 kHz and a live time of $70 \%$, for a 
luminosity of about $10^{34}$~cm$^{-2}\cdot$s$^{-1}$ and a beam energy of 
6.064~GeV.

@@

Based on the kinematics of the electrons in the calibration data, we generated 
the Helium nucleus in our RTPC GEANT4 simulation~\cite{GEANT4}. Then,
we compared the calculated GEANT4 trajectory of the Helium nuclei to 
the hits measured in the chamber. To perform this drift path extraction, 
we made a first approximation assuming a linear dependence between the radius 
of emission of the charge and its time of detection, and then refined our 
result. Indeed, because of the magnetic field, the drift paths are not linear 
in the RTPC, but the curvature was minimal and this process converged already 
on the second iteration.

@@

\begin{figure}[tb]
\centering
\includegraphics[scale=0.42]{FitResult_p2_11.pdf}
\caption{$\Delta \phi$ versus T distribution for tracks
from one bin in longitudinal position along the RTPC. The black line represents 
the final drift paths in this bin.}
\label{fig:DELTA_PHI_TDC}
\end{figure}

To verify the stability of the drift paths, this procedure was carried out 
using both the 1.204 GeV data from the beginning of the run period and the 
1.269 GeV data from the end of the run period (shown in blue on 
Figure~\ref{fig:Drift_run_number_1}). Interestingly, we found very similar 
drift paths for the two data sets and concluded that any changes in the system 
only significantly affected the drift speed.

@@

\begin{figure}[tb]
   \centering
\includegraphics[scale=0.42]{Event_ADC_Graph_8799283.pdf}
\caption{Simulated (upper) and experimental (lower) ADC and T distributions 
of a track. The colors indicate the pads, same color in top and bottom indicate 
that they are the same pad.}
\label{fig:EVENT_adc_tdc}
\end{figure}

\begin{figure*}[tp]
\centering
\includegraphics[scale=0.73]{f_dedx_p_exp_2nd.pdf}
\caption{$\small{\frac{dE}{dX}}$ vs. $p/q$ distributions for the left (on the 
   left) and for the right (on the right) half of the RTPC after gain 
   calibration. The lines are theoretical expectations from the Bethe-Bloch 
   formula for $^4$He 
   (red), $^3$He (black), $^3$H (blue), $^2$H (pink) and protons (gray).}
\label{fig:dedx_p_exp_2nd}
\end{figure*}

@@

\begin{figure}[tb!]\centering
\includegraphics[scale=0.25]{noisy_pad_before_rejection2.png}
\includegraphics[scale=0.25]{noisy_pad_after_rejection2.png}
\caption{The ADC vs. T spectrum for an example noisy channel before (top) and 
after (bottom) noise rejection algorithms.  Only hits associated with tracks 
are included, and the selection of events and tracks is the same in both 
plots.}
\label{fig:noise}
\end{figure}

>>>>

\documentclass[preprint,5p]{elsarticle}
\usepackage{graphicx}
\usepackage{dcolumn}
\usepackage{bm}  
\usepackage{amssymb}  
\usepackage{hyperref}
\providecommand{\keywords}[1]{\textbf{\textit{Keywords:}} #1}
\hypersetup{colorlinks=true, urlcolor=blue, citecolor=blue}
\usepackage[displaymath]{lineno}
\linenumbers
\usepackage{dblfloatfix}    % To enable figures at the bottom of page
\begin{document}

@@

We detail here the different regions shown in Figure~\ref{fig:RTPC_1_4} 
starting from the beam line towards larger radius:\\
\begin{itemize}
  \item The 6~atm $^4$He target extends along the beamline forming the detector 
     central axis; it is a 292~mm~long, 6~mm diameter Kapton straw with a 
     27~$\mu$m wall. The detector and the target are centered at 64~cm upstream 
     of the CLAS center.
   \item The first gas gap covers the radial range from 3~mm to 20~mm. It is 
      filled with $^{4}$He gas at 1~atm to minimize secondary interactions from
      M\o{}ller electrons scattered by the beam. This region is surrounded by 
      a 4 $\mu$m thick window made of grounded aluminized Mylar.
   \item The second gas gap region extends between 20~mm and 30~mm and is 
      filled with the gas mixture of 80$\%$ neon (Ne) and 20$\%$ dimethyl ether 
      (DME).  This region is surrounded by a 4~$\mu$m thick window made of 
      aluminized Mylar set at $-4260$~V to serve as the cathode.
   \item The drift region is filled with the same Ne-DME gas mixture and 
      extends from the cathode to the first GEM, 60~mm away from the beam axis.  
      The average electric field in this region is perpendicular to the beam 
      and about 550~V/cm.
   \item The electron amplification system is composed of three GEMs located at 
      radii of 60, 63 and 66~mm. In this configuration, the first GEM layer 
      serves as the anode and each subsequent GEM is set at a lower voltage to
      obtain a strong ($\sim$1600~V/cm) electric field between the GEM foils. A 
      275~V bias is applied across each GEM for amplification.
   \item The readout board has an internal radius of 69~mm and collects charges 
      after they have been multiplied by the GEMs. Pre-amplifiers are plugged 
      directly on its outer side and transmit the signal to the data 
      acquisition electronics.
\end{itemize}

@@

The trigger signal also initiates RCU readout operation from FEC boards. All the 
measured samples from active channels are reported, together with a channel 
identifier and a timestamp, to the ROC application, that in turn sends them on 
the main event builder. Readout from FECs occurs in ``block-transfer'' mode, 
making use of the ALTRO internal multi-event buffer. This feature could be 
exploited thanks to the new RCU boards developed for the RTPC readout 
system\footnote{In the original BoNuS system, the U2F readout controller had no 
   multi-event capability and used a USB connection rather than fiber optics 
   for readout. These factors limited the maximum readout rate from the BoNuS 
detector to about 500 Hz.}. In order to read all the detector readout pads, 
four FEC crates were used, each equipped with 6 boards, plus a ROC. A schematic 
of the readout system, for a single crate, is reported in 
Figure~\ref{fig:ReadoutScheme}. This configuration permitted to reduce the dead 
time associated to FEC readout operations, which is scaling linearly with the 
number of boards in the crate. During the 2009 run, the system was successfully 
operated with a DAQ rate of 3.1 kHz and a live time of $70 \%$, for a 
luminosity of about $10^{34}$~cm$^{-2}\cdot$s$^{-1}$ and a beam energy of 
6.064~GeV.

@@

Based on the kinematics of the electrons in the calibration data, we generated 
the Helium nucleus in our RTPC GEANT4 simulation~\cite{GEANT4}. Then,
we compared the calculated GEANT4 trajectory of the Helium nuclei to 
the hits measured in the chamber. To perform this drift path extraction, 
we made a first approximation assuming a linear dependence between the radius 
of emission of the charge and its time of detection, and then refined our 
result. Because of the magnetic field, the drift paths in RTPC are not linear 
but the curvature is small and the refining  process converged already on the 
second iteration.

@@

\begin{figure}[tb]
\centering
\includegraphics[scale=0.46]{FitResult_p2_11.pdf}
\caption{$\Delta \phi$ versus T distribution for tracks
from one bin in longitudinal position along the RTPC. The black line represents 
the final drift paths in this bin.}
\label{fig:DELTA_PHI_TDC}
\end{figure}

To verify the stability of the drift paths, this procedure was carried out 
using both the 1.204 GeV data from the beginning of the run period and the 
1.269 GeV data from the end of the run period (shown in blue on 
Figure~\ref{fig:Drift_run_number_1}). We found very similar drift paths for the 
two data sets and concluded that any changes in the system only significantly 
affected the drift speed.

@@

\begin{figure}[!t]
   \centering
\includegraphics[scale=0.42]{Event_ADC_Graph_8799283.pdf}
\caption{Simulated (upper) and experimental (lower) ADC and T distributions 
of a track. The colors indicate the pads, same color in top and bottom indicate 
that they are the same pad.}
\label{fig:EVENT_adc_tdc}
\end{figure}

\begin{figure*}[!h]
\centering
\includegraphics[scale=0.73]{f_dedx_p_exp_2nd.pdf}
\caption{$\small{\frac{dE}{dX}}$ vs. $p/q$ distributions for the left (on the 
   left) and for the right (on the right) half of the RTPC after gain 
   calibration. The lines are theoretical expectations from the Bethe-Bloch 
   formula for $^4$He 
   (red), $^3$He (black), $^3$H (blue), $^2$H (pink) and protons (gray).}
\label{fig:dedx_p_exp_2nd}
\end{figure*}

@@

\begin{figure}[tb]
   \centering
\includegraphics[scale=0.25]{noisy_pad_before_rejection2.png}
\includegraphics[scale=0.25]{noisy_pad_after_rejection2.png}
\caption{The ADC vs. T spectrum for an example noisy channel before (top) and 
after (bottom) noise rejection algorithms.  Only hits associated with tracks 
are included, and the selection of events and tracks is the same in both 
plots.}
\label{fig:noise}
\end{figure}
